---
title: Package downloads analysis 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(rvest)
library(glue)
library(feasts)
library(cranlogs)
library(plotly)
library(scales)
library(lubridate)

```

remove missing data and replace it with the closest date figure 
```{r total_download}
load(here::here("data/derived/total_downloads.RData"))

# replace missing values with the nearest number 
total_downloads$count[total_downloads$count == 0] <- 21959

total_downloads %>%
  ggplot() + geom_line(aes(date, count))+
  geom_smooth(aes(date, count),stat = "smooth") 

```

Does R related conferences affect the total R packages downloads?

```{r conferences-dates}
# adding the userR conference date
conference_dates <- tribble(
  ~name, ~date,
  "UseR", "2004-05-20",
  "UseR", "2006-06-15",
  "UseR", "2007-08-08",
  "UseR", "2008-08-14",
  "UseR", "2009-07-08",
  "UseR", "2010-07-20",
  "UseR", "2011-08-16",
  "UseR", "2012-06-12",
  "UseR", "2013-06-10",
  "UseR", "2014-06-30",
  "UseR", "2015-06-30",
  "UseR", "2016-06-27",
  "UseR", "2017-07-04",
  "UseR", "2018-07-10",
  "UseR", "2019-07-09",
  "UseR", "2020-07-07",
  "UseR", "2021-07-05",
  "RStudio", "2017-01-11",
  "RStudio", "2018-01-31",
  "RStudio", "2019-01-15",
  "RStudio", "2020-01-27",
  "RStudio", "2021-03-03",
)

UseR <- dplyr::filter(conference_dates, name == "UseR", date > "2013-01-01")
Rstudio <- dplyr::filter(conference_dates, name == "RStudio")

download_user <- total_downloads %>%
  ggplot() +
  geom_line(aes(date, count)) +
  geom_vline(
    xintercept = as.numeric(as.Date(UseR$date)), linetype = "dotted",
    color = "blue", size = 0.1
  ) +
  geom_vline(
    xintercept = as.numeric(as.Date(Rstudio$date)), linetype = "dotted",
    color = "red", size = 0.1
  )

ggplotly(download_user, dynamicTicks = TRUE) %>%
  layout(hovermode = "x")
```

Does working days have an effect on the R packages downloads?

```{r working-day-effect}
# working days effect 
total_downloads %>% 
  as_tsibble(index = date) %>% 
  gg_subseries(count,period = "week")
```

Does American academic calendar effect number of R packages downloads? 

```{r}

```


```{r STL}
# STL decomposition of total R package downloads
total_downloads %>%
  as_tsibble(index = date) %>%
  model(
    STL(log(count + 1) ~ trend(window = 77) +
      season("week", window = "periodic"))
  ) %>%
  components() %>%
  autoplot() +
  geom_vline(xintercept = as.numeric(as.Date(UseR$date)), linetype = "dotted", color = "blue", size = 0.1) +
  geom_vline(xintercept = as.numeric(as.Date(Rstudio$date)), linetype = "dotted", color = "red", size = 0.1)

```




R packages download and updates analysis: 

```{r updates-data}
load(here::here("data/derived/updates.Rdata"))
load(here::here("data/derived/updates_2.Rdata"))
load(here::here("data/derived/updates_3.Rdata"))
load(here::here("data/derived/updates_4.Rdata"))
load(here::here("data/derived/updates_5.Rdata"))
load(here::here("data/derived/updates_6.Rdata"))
load(here::here("data/derived/updates_7.Rdata"))
load(here::here("data/derived/updates_8.Rdata"))
load(here::here("data/derived/updates_9.Rdata"))

updates_list <- list(updates, updates_2, updates_3, updates_4, updates_5, updates_6, updates_7, updates_8, updates_9)
updates_select <- map_dfr(updates_list, select_pkg_ver_date)

updates_select <- updates_select %>%
  mutate(Date = as.Date(updates_select$date))
```


```{r }
### No. of packages on CRAN before RStudio came along
nrow(updates_select %>%
  group_by(Package) %>%
  filter(Date <= "2016-11-01") %>%
  filter(Version == max(Version)))


### No. of packages on CRAN after RStudio came along

nrow(updates_select %>%
  group_by(Package) %>%
  filter(Date >= "2016-11-01") %>%
  filter(Version == max(Version)))

### Updates summaries statistics 
update_freq <- updates_select %>% 
  group_by(Package) %>% 
  arrange(Date) %>%
  mutate(diff = c(NA,diff(Date)))

update_avg <- update_freq %>% 
  select(Package,diff) %>% 
  na.omit() %>% 
  group_by(Package) %>%
  summarise(avg=mean(diff))

 updates_select%>% 
  filter(Package=="Ecdat")

summary(update_freq$diff)
```

```{r}
# Select top 10 packages from last month 

top_downloads <- cran_top_downloads("last-month")

updates_top <- updates_select %>%
  filter(Package %in% top_downloads$package) %>%
  filter(Date >= as.Date("2013-01-01"))


downloads_top <- cran_downloads(from = "2013-01-01", to = "2021-07-05", package = top_downloads$package)

downloads_top %>%
  ggplot() +
  geom_line(aes(date, count, col = package))
  

```


```{r}
# updates and ggplot2 downloads pattern 

ggplot2 <- downloads_top %>%
  filter(package == "ggplot2") %>%
  ggplot() +
  geom_line(aes(date, count)) +
  geom_vline(
    xintercept = updates_top$Date, linetype = "dotted", color = "red",
    size = 0.1
  )

ggplotly(ggplot2, dynamicTicks = TRUE) %>%
  layout(hovermode = "x")
```
```{r}
# STL decomposition on ggplot 
STL_ggplot <- downloads_top %>%
  filter(package == "ggplot2") %>%
  as_tsibble(index = date) %>%
  model(
    STL(log(count + 1) ~ trend(window = NULL) +
      season("week", window = "periodic"))
  ) %>%
  components() %>%
  autoplot() +
  geom_vline(xintercept = updates_top$Date, linetype = "dotted", color = "blue", size = 0.1)

ggplotly(STL_ggplot, dynamicTicks = TRUE) %>%
  layout(hovermode = "x")
```

```{r}
# updates and dplyr downloads pattern 

dplyr <- downloads_top %>%
  filter(package == "dplyr") %>%
  ggplot() +
  geom_line(aes(date, count)) +
  geom_vline(
    xintercept = updates_top$Date, linetype = "dotted", color = "red",
    size = 0.1
  )

ggplotly(dplyr, dynamicTicks = TRUE) %>%
  layout(hovermode = "x")
```




```{r}
# non-RStudio packages updates

updates_forecast <- updates_select %>%
  filter(Package == "forecast") %>%
  filter(Date >= as.Date("2012-10-29"))


forecast_download <- cran_downloads(packages = "forecast", from = "2012-10-29", to = "2021-07-12")

forecast <- forecast_download %>%
  ggplot() +
  geom_line(aes(date, count)) +
  geom_vline(xintercept = updates_forecast$Date, linetype = "dotted", color = "red", size = 0.1) +
  scale_x_date(limits = as.Date(c("2012-10-29", "2021-07-12")))

ggplotly(forecast, dynamicTicks = TRUE) %>%
  layout(hovermode = "x")

```


```{r}
#STL decompostion to forecast package downloads 
STL_forecast <- forecast_download %>%
  as_tsibble(index = date) %>%
  model(
    STL(log(count + 1) ~ trend(window = NULL) +
      season("week", window = "periodic"))
  ) %>%
  components() %>%
  autoplot() +
  geom_vline(xintercept = updates_forecast$Date, linetype = "dotted", color = "blue", size = 0.1)

ggplotly(STL_forecast, dynamicTicks = TRUE) %>%
  layout(hovermode = "x")

```



Number of pacakges being released per day: 

```{r}
first_release <- updates_select %>% 
  group_by(Package) %>% 
  filter(Date == min(Date))

first_release %>%
  group_by(Date) %>%
  summarise(n = n_distinct(Package)) %>%
  ggplot()  + geom_line(aes(Date, n))+
  geom_smooth(aes(Date, n),stat = "smooth")


```


```{r}

#getting the top 100 downloads for last month 

download_top100 <- get_downloads(when = "last-month")

top_100 <- db %>%
  mutate(Package = as.character(Package)) %>%
  filter(Package %in% download_top100$package) %>%
  filter(!grepl("RStudio", Author)) %>%
  select(Package)

# Non RStudio downloads
top_nonrstudio <- download_top100 %>%
  filter(package %in% top_100$Package)


updates_jsonlite <- updates_select %>%
  filter(Package == "jsonlite") %>%
  filter(Date >= as.Date("2012-10-29"))

# Non RStudio downloads
top_nonrstudio <- download_top100 %>%
  filter(package %in% top_100$Package)


```

package updates effect to non-RStudio own popular packages 

```{r }
# package updates on jsonlite
updates_jsonlite <- updates_select %>%
  filter(Package == "jsonlite") %>%
  filter(Date >= as.Date("2012-10-29"))


jsonlite_download <- cran_downloads(packages = "jsonlite", from = "2013-12-04", to = "2021-07-12")

jsonlite <- jsonlite_download %>%
  ggplot() +
  geom_line(aes(date, count)) +
  geom_vline(xintercept = updates_jsonlite$Date, linetype = "dotted", color = "red", size = 0.1) +
  scale_x_date(limits = as.Date(c("2013-12-04", "2021-07-12")))

ggplotly(jsonlite, dynamicTicks = TRUE) %>%
  layout(hovermode = "x")

```


```{r}
# package updates on glue
updates_glue <- updates_select %>%
  filter(Package == "glue") %>%
  filter(Date >= as.Date("2012-10-29"))


glue_download <- cran_downloads(packages = "glue", from = "2017-04-17", to = "2021-07-12")

glue <- glue_download %>%
  ggplot() +
  geom_line(aes(date, count)) +
  geom_vline(xintercept = updates_glue$Date, linetype = "dotted", color = "red", size = 0.1) +
  scale_x_date(limits = as.Date(c("2017-04-17", "2021-07-12")))

ggplotly(glue, dynamicTicks = TRUE) %>%
  layout(hovermode = "x")
```

Selecting packages that have more than 1 updates 

```{r}
updates_select %>%
  group_by(Package) %>%
  summarise(n = n(), last = max(Date)) %>%
  filter(n > 2) %>%
  arrange(desc(n))
```


How to get the package updates (Alternative way):

```{r pkg-updates}
pkg_url <- "https://cran.r-project.org/web/packages/{pkg}/index.html"
pkg_archive <- "https://cran.r-project.org/src/contrib/Archive/{pkg}/"

pkgs_of_interest <- doe_pkgs[1:4]
pkg_updates <- map(pkgs_of_interest, function(pkg) {
    last_update <- read_html(glue(pkg_url)) %>% 
      html_table() %>% 
      .[[1]] %>% 
      filter(X1=="Published:") %>% 
      pull(X2) %>% 
      ymd()
      
    archive_dates <- tryCatch({ 
        read_html(glue(pkg_archive)) %>% 
          html_table() %>%
          .[[1]] %>% 
          pull(`Last modified`) %>% 
          ymd_hm() %>% 
          na.omit() %>% 
          as.Date()
      }, error = function(e) {
        NULL
      })
    c(archive_dates, last_update)
  })
names(pkg_updates) <- pkgs_of_interest

updates <- unlist(pkg_updates) %>% 
  enframe("package", "update") %>% 
  # unlist converts date to integers
  mutate(update = as.Date(update, origin = "1970-01-01"),
         # need to get rid of the numbers appended to pkg names
         package = str_extract(package, paste0(pkgs_of_interest, collapse="|"))) 

updates
```

